---
title: "Normalization for `r params$dataset`"
author: "Lei Jiang"
date: "`r Sys.Date()`"
params:
  unique_id: normalization # unique_id for live log, which could be job_id, user_id or process_id
  dataset: NA
  input: NULL # Input file path to scRNA data.
  output: NULL # Output file path of the Seurat object.
  adata_path: NULL
  methods: NULL # Normalization methods: c("logCP10K", "logCPM", "SCT", "SCT_V2", 'sctransform', 'sctransform_v2', "scran", "DESeq2", "RLE", "TMM", "upperquartile", "UQ", "TPM", "FPKM", "Pearson_residuals")
  default_assay: RNA # Default assay for each normalization method:  "RNA", "MAGIC_imputed", "SAVER_imputed", "scGNN_imputed"
  output_format: AnnData # Seurat or AnnData
  species: NULL # c("human", "mouse") Species of the database for annotation. Allowed input is human or mouse.
  idtype: ENSEMBL # idtype should be one of "SYMBOL", "ENSEMBL", "ENTREZID" or "REFSEQ".
  show_umap: FALSE
  show_code: FALSE
  show_warning: FALSE
  show_message: TRUE
  show_error: TRUE
output: html_document
---
```{r setup, echo=params$show_code, include=FALSE, warning=params$show_warning, message=params$show_message}
knitr::opts_chunk$set(echo=params$show_code, include=TRUE, warning=params$show_warning, message=params$show_message)
```
# Normalization
To accurately normalize the gene expression values to account for differences in sequencing depth and overdispersed count values.
```{r Load data, echo=params$show_code, warning=params$show_warning, message=params$show_message}
if(is.null(params$methods)) stop("No normalization method is selected.")
if(is.null(params$input)) stop("Inupt_path can not be NULL.")
if(is.null(params$output)) stop("Output_path can not be NULL.")
source("../formating/formating.R")
source("../visualization/visualization.R")
srat <- NULL
adata <- NULL
unique_id <- params$unique_id
# Check if output file already exists
if(file.exists(params$output)){
  srat <- tryCatch(
    LoadSeurat(params$output, raw=TRUE),
    error = function(e) {
        stop("The file format is not supported.")
        if(params$show_error) print(e)
    }
  )
}else{
  srat <- tryCatch(
    LoadSeurat(params$input, raw=TRUE),
    error = function(e) {
        stop("The file format is not supported.")
        if(params$show_error) print(e)
    }
  )
}
if(is.null(srat)) stop("The file format is not supported.")
if(!params$default_assay %in% names(srat@assays)) stop(paste(params$default_assay, "does not exist."))
DefaultAssay(object = srat) <- params$default_assay
print(paste("Seurat object is loaded and default assay is set to", params$default_assay, "."))

adata_path <- params$adata_path
if (file.exists(adata_path)){
  adata <- LoadAnndata(adata_path)
} else{
  adata <- SeuratToAnndata(srat, assay=params$default_assay)
}

expression_matrix <- srat[[params$default_assay]]@counts
if(IsNormalized(expression_matrix)){
  RedisError(unique_id, "Normalization only takes raw counts, not normalized data.")
  stop("Normalization only takes raw counts, not normalized data.")
}

adata$layers["raw_counts"] <- adata$X
print(srat)
# methods <- eval(parse(text = params$methods))
methods <- strsplit(toupper(params$methods), ",")[[1]]
```
## logCP10K
Scale data to account for sequencing depth by 10,000 (as if all cells have 10k UMIs overall), and log2-transform the obtained values.
```{r logCP10K, error=TRUE, tidy=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
if("LOGCP10K" %in% methods){
  if(!"logCP10K" %in% names(srat@assays)){
    tryCatch({
      RedisInfo(unique_id, "Running logCP10K normalization.")
      srat[['logCP10K']] <- CreateAssayObject(counts = expression_matrix)
      srat <- NormalizeData(srat, assay = 'logCP10K')
      SaveSeurat(srat, params$output)
      RedisInfo(unique_id, "Seurat object for logCP10K is saved successfully.")
      
      adata$layers["logCP10K"] <- t(as.data.frame(srat[['logCP10K']]@data))
      RedisInfo(unique_id, "logCP10K layer has been added to AnnData object.")
      write_h5ad(adata, adata_path, compression = "gzip")
      RedisInfo(unique_id, "AnnData object for logCP10K is saved successfully.")
    }, error = function(e) {
      RedisWarning(unique_id, paste0("An error happened during logCP10K normalization, skipped: ", e$message))
    }) 
  } else {
    RedisInfo(unique_id, "logCP10K assay already exists, skipped.")
    if(!"logCP10K" %in% names(adata$layers)){
      adata$layers["logCP10K"] <- t(as.data.frame(srat[['logCP10K']]@data))
      RedisInfo(unique_id, "logCP10K layer has been added to AnnData object.")
      write_h5ad(adata, adata_path, compression = "gzip")
      RedisInfo(unique_id, "AnnData object for logCP10K is saved successfully.")
    }
  }
  # knitr::kable(head(srat[['logCP10K']]@data), align='cccc', format = "html", caption = 'logCP10K')
  if(params$show_umap) seurat_cluster_umap(srat, assay = 'logCP10K', title = "logCP10K normalization")
} else {
  print("Not selected by user, skipped.")
}
```

## logCPM
Load or create a SingleCellExperiment object from a expression matrix where each row corresponds to a gene (or transcript) and each column corresponds to a single cell.
```{r logCPM, error=TRUE, tidy=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
if("LOGCPM"  %in% methods){
  if(!"logCPM" %in% names(srat@assays)){  
    tryCatch({
      RedisInfo(unique_id, "Running logCPM normalization.")
      srat[['logCPM']] <- CreateAssayObject(counts = expression_matrix)
      srat <- NormalizeData(srat, scale.factor = 1000000, assay = 'logCPM')
      SaveSeurat(srat, params$output)
      RedisInfo(unique_id, "Seurat object for logCPM is saved successfully.")

      adata$layers["logCPM"] <- t(as.data.frame(srat[['logCPM']]@data))
      RedisInfo(unique_id, "logCPM layer has been added to AnnData object.")
      write_h5ad(adata, adata_path, compression = "gzip")
      RedisInfo(unique_id, "AnnData object for logCPM is saved successfully.")
    }, error = function(e) {
      RedisWarning(unique_id, paste0("An error happened during logCPM normalization, skipped: ", e$message))
    }) 
  } else {
    RedisInfo(unique_id, "logCPM assay already exists, skipped.")
    if(!"logCPM" %in% names(adata$layers)){
      adata$layers["logCPM"] <- t(as.data.frame(srat[['logCPM']]@data))
      RedisInfo(unique_id, "logCPM layer has been added to AnnData object.")
      write_h5ad(adata, adata_path, compression = "gzip")
      RedisInfo(unique_id, "AnnData object for logCPM is saved successfully.")
    }
  }
  # knitr::kable(head(srat[['logCPM']]@data), align='cccc', format = "html", caption = 'logCPM')
  if(params$show_umap) seurat_cluster_umap(srat, assay = 'logCPM', title = "logCPM normalization")
} else {
  print("Not selected by user, skipped.")
}
```

## SCTransform
SCTransform calculates a model of technical noise in scRNA-seq data using 'regularized negative binomial regression'. The residuals for this model are normalized values, and can be positive or negative. Positive residuals for a given gene in a given cell indicate that we observed more UMIs than expected given the gene’s average expression in the population and cellular sequencing depth, while negative residuals indicate the converse.
```{r SCTransform, error=TRUE, tidy=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
if(("SCTRANSFORM" %in% methods) || ("SCTRANSFORM_V2" %in% methods) || ("SCT" %in% methods) || ("SCT_V2" %in% methods)){
  
  if(!"SCT" %in% names(srat@assays)){
    tryCatch({
      library(sctransform)
      if(("SCTRANSFORM"  %in% methods) || ("SCT" %in% methods)){
        RedisInfo(unique_id, "Running SCTransform normalization.")
        # store mitochondrial percentage in object meta data
        srat <- PercentageFeatureSet(srat, pattern = "^MT-", col.name = "percent.mt")
        # srat <- Seurat::SCTransform(srat, method = "glmGamPoi", vars.to.regress = params$vars.to.regress, verbose = FALSE)
        srat <- Seurat::SCTransform(srat, vars.to.regress = "percent.mt", verbose = FALSE)
        DefaultAssay(object = srat) <- params$default_assay
        SaveSeurat(srat, params$output)
        RedisInfo(unique_id, "Seurat object for SCTransform is saved successfully.")

        adata_sct <- SeuratToAnndata(srat, assay="SCT")
        adata_sct$layers["SCTransform"] <- adata_sct$X
        write_h5ad(adata_sct, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for SCTransform is saved separately successfully.")
        adata_sct <- NULL
      }
      else{
        RedisInfo(unique_id, "Running SCTransform VST V2 normalization.")
        srat <- Seurat::SCTransform(srat, vst.flavor="v2", verbose = FALSE)
        DefaultAssay(object = srat) <- params$default_assay
        SaveSeurat(srat, params$output)
        RedisInfo(unique_id, "Seurat object for SCTransform VST V2 is saved successfully.")
        adata_sct <- SeuratToAnndata(srat, assay="SCT")
        adata_sct$layers["SCTransform_v2"] <- adata_sct$X
        write_h5ad(adata_sct, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for SCTransform is saved separately successfully.")
        adata_sct <- NULL
      }

      # adata_sct <- AnnData(X = t(srat[['SCTransform']]@data), obs = srat@meta.data, var = srat[['SCTransform']]@meta.features)
      # adata_sct.path <-  gsub(".h5ad", "_SCT.h5ad", adata_path)
      # adata_sct <- SeuratToAnndata(srat, out_file=adata_sct.path, assay="SCT")
    }, error = function(e) {
      RedisWarning(unique_id, paste0("An error happened during SCTransform normalization, skipped: ", e$message))
    })
  } else {
    RedisInfo(unique_id, "SCTransform assay already exists, skipped.")
    # adata_sct <- AnnData(X = t(srat[['SCTransform']]@data), obs = srat@meta.data, var = srat[['SCTransform']]@meta.features)
    # adata_sct.path <-  gsub(".h5ad", "_SCT.h5ad", adata_path)
    # adata_sct <- SeuratToAnndata(srat, out_file=adata_sct.path, assay="SCT")
    adata_sct <- SeuratToAnndata(srat, assay="SCT")
    adata_sct$layers["SCTransform"] <- adata_sct$X
    write_h5ad(adata_sct, adata_path, compression = "gzip")
    RedisInfo(unique_id, "AnnData object for SCTransform is saved separately successfully.")
    adata_sct <- NULL
  }
  # knitr::kable(head(srat[['SCT']]@data), align='cccc', format = "html", caption = 'SCTransform')
  if(params$show_umap) seurat_cluster_umap(srat, assay = 'SCT', title = "SCTransform normalization")
} else {
  print("Not selected by user, skipped.")
}
```

## scran
scran deals with the problem of vary large numbers of zero values per cell by pooling cells together calculating a normalization factor (similar to CPM) for the sum of each pool. Since each cell is found in many different pools, cell-specific factors can be deconvoluted from the collection of pool-specific factors using linear algebraic methods.
```{r scran, error=TRUE, tidy=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
if("SCRAN" %in% methods){
  if(!"scran" %in% names(srat@assays)){
    tryCatch({
      RedisInfo(unique_id, "Running scran normalization.")
      library(scRNA.seq.funcs)
      library(scran)
      set.seed(1234567)
      qclust <- quickCluster(expression_matrix, min.size = 30)
      table(qclust)
      sce <- SingleCellExperiment(assays = list(counts = expression_matrix))
      sce <- computeSumFactors(sce, clusters = qclust)
      sce <- logNormCounts(sce)
      sran_assay <- CreateAssayObject(data = logcounts(sce))
      srat[['scran']] <- sran_assay
      sce <- NULL
      SaveSeurat(srat, params$output)
      RedisInfo(unique_id, "Seurat object for sran is saved successfully.")

      adata$layers["scran"] <- t(as.data.frame(srat[['scran']]@data))
      RedisInfo(unique_id, "scran layer has been added to AnnData object.")
      write_h5ad(adata, adata_path, compression = "gzip")
      RedisInfo(unique_id, "AnnData object for scran is saved successfully.")
    }, error = function(e) {
      RedisWarning(unique_id, paste0("An error happened during scran normalization, skipped: ", e$message))
    }) 
  } else {
    RedisInfo(unique_id, "scran assay already exists, skipped.")
    if(!"scran" %in% names(adata$layers)){
      adata$layers["scran"] <- t(as.data.frame(srat[['scran']]@data))
      RedisInfo(unique_id, "scran layer has been added to AnnData object.")
      write_h5ad(adata, adata_path, compression = "gzip")
      RedisInfo(unique_id, "AnnData object for scran is saved successfully.")
    }
  }
  # knitr::kable(head(srat[['scran']]@data), align='cccc', format = "html", caption = 'scran')
  if(params$show_umap) seurat_cluster_umap(srat, assay = 'scran', title = "scran normalization")
} else {
  print("Not selected by user, skipped.")
}
```

## DESeq2 (RLE)
RLE (Relative Log Expression) 
```{r DESeq2, error=TRUE, tidy=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
if("DESEQ2" %in% methods){
  if(!"DESeq2" %in% names(srat@assays)){
    tryCatch({
      RedisInfo(unique_id, "Running DESeq2 normalization.")
      library( "DESeq2" )
      library( "SummarizedExperiment" )
      se <- SummarizedExperiment(as.matrix(expression_matrix + 1))
      dds <- DESeqDataSet(se, design=~1)
      dds <- estimateSizeFactors(dds)
      dds_assay <- CreateAssayObject(data = counts(dds, normalized=TRUE))
      srat[['DESeq2']] <- dds_assay
      SaveSeurat(srat, params$output)
      RedisInfo(unique_id, "Seurat object for DESeq2 is saved successfully.")

      adata$layers["DESeq2"] <- t(as.data.frame(srat[['DESeq2']]@data))
      RedisInfo(unique_id, "DESeq2 layer has been added to AnnData object.")
      write_h5ad(adata, adata_path, compression = "gzip")
      RedisInfo(unique_id, "AnnData object for DESeq2 is saved successfully.")
    }, error = function(e) {
      RedisWarning(unique_id, paste0("An error happened during DESeq2 normalization, skipped: ", e$message))
    }) 
  } else {
    RedisInfo(unique_id, "DESeq2 assay already exists, skipped.")
    if(!"DESeq2" %in% names(adata$layers)){
      adata$layers["DESeq2"] <- t(as.data.frame(srat[['DESeq2']]@data))
      RedisInfo(unique_id, "DESeq2 layer has been added to AnnData object.")
      write_h5ad(adata, adata_path, compression = "gzip")
      RedisInfo(unique_id, "AnnData object for DESeq2 is saved successfully.")
    }
  }
  # knitr::kable(head(srat[['DESeq2']]@data), align='cccc', format = "html", caption = 'DESeq2')
  if(params$show_umap) seurat_cluster_umap(srat, assay = 'DESeq2', title = "DESeq2 normalization")
} else {
  print("Not selected by user, skipped.")
}
```

## edgeR (TMM, RLE, UpperQuartile (UQ))
```{r edgeR, error=TRUE, tidy=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
if("TMM" %in% methods || "RLE" %in% methods || "UQ" %in% methods || "UPPERQUARTILE" %in% methods){
  library(edgeR)
  dge <- DGEList(expression_matrix + 1)
  if("TMM" %in% methods){
    if(!"TMM" %in% names(srat@assays)){    
      tryCatch({
        RedisInfo(unique_id, "Running TMM normalization.")
        tmm <- calcNormFactors(dge, method="TMM")
        tmm_assay <- CreateAssayObject(data = edgeR::cpm(tmm))
        srat[['TMM']] <- tmm_assay
        SaveSeurat(srat, params$output)
        RedisInfo(unique_id, "Seurat object for TMM is saved successfully.")

        adata$layers["TMM"] <- t(as.data.frame(srat[['TMM']]@data))
        RedisInfo(unique_id, "TMM layer has been added to AnnData object.")
        write_h5ad(adata, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for TMM is saved successfully.")
      }, error = function(e) {
        RedisWarning(unique_id, paste0("An error happened during TMM normalization, skipped: ", e$message))
      }) 
    } else {
      RedisInfo(unique_id, "TMM assay already exists, skipped.")
      if(!"TMM" %in% names(adata$layers)){
        adata$layers["TMM"] <- t(as.data.frame(srat[['TMM']]@data))
        RedisInfo(unique_id, "TMM layer has been added to AnnData object.")
        write_h5ad(adata, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for TMM is saved successfully.")
      }
    }
    # knitr::kable(head(srat[['TMM']]@data), align='cccc', format = "html", caption = 'TMM')
    if(params$show_umap) seurat_cluster_umap(srat, assay = 'TMM', title = "TMM normalization")
  } else {
    print("TMM is not selected by user, skipped.")
  }

  if("RLE" %in% methods){
    if(!"RLE" %in% names(srat@assays)){
      tryCatch({
        RedisInfo(unique_id, "Running RLE normalization.")
        rle <- calcNormFactors(dge, method="RLE")
        rle_assay <- CreateAssayObject(data = edgeR::cpm(rle))
        srat[['RLE']] <- rle_assay
        SaveSeurat(srat, params$output)
        RedisWarning(unique_id, "Seurat object for RLE is saved successfully.")

        adata$layers["RLE"] <- t(as.data.frame(srat[['RLE']]@data))
        RedisInfo(unique_id, "RLE layer has been added to AnnData object.")
        write_h5ad(adata, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for RLE is saved successfully.")
      }, error = function(e) {
        RedisWarning(unique_id, paste0("An error happened during RLE normalization, skipped: ", e$message))
      }) 
    } else {
      RedisInfo(unique_id, "RLE assay already exists, skipped.")
      if(!"RLE" %in% names(adata$layers)){
        adata$layers["RLE"] <- t(as.data.frame(srat[['RLE']]@data))
        RedisInfo(unique_id, "RLE layer has been added to AnnData object.")
        write_h5ad(adata, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for RLE is saved successfully.")
      }
    }
    # knitr::kable(head(srat[['RLE']]@data), align='cccc', format = "html", caption = 'RLE')
    if(params$show_umap) seurat_cluster_umap(srat, assay = 'RLE', title = "RLE normalization")
    } else {
      print("RLE is not selected by user, skipped.")
    }

  if("UQ" %in% methods || "UPPERQUARTILE" %in% methods){
    RedisInfo(unique_id, "Running Upper Quartile normalization.")
    if(!"UpperQuartile" %in% names(srat@assays)){
      tryCatch({
        uq <- calcNormFactors(dge, method="upperquartile")
        uq_assay <- CreateAssayObject(data = edgeR::cpm(uq))
        srat[['UpperQuartile']] <- uq_assay
        SaveSeurat(srat, params$output)
        RedisInfo(unique_id, "Seurat object for UpperQuartile is saved successfully.")

        adata$layers["UpperQuartile"] <- t(as.data.frame(srat[['UpperQuartile']]@data))
        RedisInfo(unique_id, "UpperQuartile layer has been added to AnnData object.")
        write_h5ad(adata, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for UpperQuartile is saved successfully.")
      }, error = function(e) {
        RedisWarning(unique_id, paste0("An error happened during UpperQuartile normalization, skipped: ", e$message))
      }) 
    } else {
      RedisInfo(unique_id, "UpperQuartile assay already exists, skipped.")
      if(!"UpperQuartile" %in% names(adata$layers)){
        adata$layers["UpperQuartile"] <- t(as.data.frame(srat[['UpperQuartile']]@data))
        RedisInfo(unique_id, "UpperQuartile layer has been added to AnnData object.")
        write_h5ad(adata, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for UpperQuartile is saved successfully.")
      }
    }
    # knitr::kable(head(srat[['UpperQuartile']]@data), align='cccc', format = "html", caption = 'UpperQuartile')
    if(params$show_umap) seurat_cluster_umap(srat, assay = 'UpperQuartile', title = "Upper quartile normalization")
  } else {
    print("UpperQuartile is not selected by user, skipped.")
  }
}else {
  print("Not selected by user, skipped.")
}
```

## scanpy Pearson residuals normalization
```{r scanpy, error=TRUE, tidy=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
if(toupper(params$output_format) == "ANNDATA" || "PEARSON_RESIDUALS" %in% methods){
    if("PEARSON_RESIDUALS" %in% methods){
      if(!"Pearson_residuals" %in% names(srat@assays)){
        library(reticulate)
        source_python("./scanpy_r.py")
        source("../formating/formating.R")
        # adata <- LoadAnndata(adata_path)
        tryCatch(
          {
            RedisInfo(unique_id, "Running Pearson_residuals normalization.")
            adata <- py_to_r_ifneedbe(normalize_pearson_residuals(r_to_py(adata, convert = FALSE))) 
            pr_assay <- CreateAssayObject(data = t(adata$layers["Pearson_residuals"]))
            RedisInfo(unique_id, "Adding Pearson_residuals assay to Seurat object.")
            srat[['Pearson_residuals']] <- pr_assay
            SaveSeurat(srat, params$output)
            RedisInfo(unique_id, "Seurat object for Pearson_residuals is saved successfully.")
            write_h5ad(adata, adata_path, compression = "gzip")
            RedisInfo(unique_id, "AnnData object for Pearson_residuals is saved successfully.")
          },
          error = function(e) {
            RedisWarning(unique_id, paste0("An error happened during Pearson residuals normalization, skipped: ", e$message))
          }
        )
      } else {
        RedisInfo(unique_id, "Pearson_residuals assay already exists, skipped.")
        if(!"Pearson_residuals" %in% names(adata$layers)){
          adata$layers["Pearson_residuals"] <- t(as.data.frame(srat[['Pearson_residuals']]@data))
          RedisInfo(unique_id, "Pearson_residuals layer has been added to AnnData object.")
          write_h5ad(adata, adata_path, compression = "gzip")
          RedisInfo(unique_id, "AnnData object for Pearson_residuals is saved successfully.")
        }
      }
      # knitr::kable(head(srat[['Pearson_residuals']]@data), align='cccc', format = "html", caption = 'Pearson_residuals')
      if(params$show_umap) seurat_cluster_umap(srat, assay = 'Pearson_residuals', title = "Pearson residuals normalization")
    } else {
      print("Not selected by user, skipped.")
    }
  }
```

## biomaRt (TPM, FPKM)
TPM (Transcripts Per Kilobase Million) is the same as RPKM/FPKM, but it divides the read counts by the gene length (in kilobase) first, then divides the total counts by 106 to make the scaling factor. In essence, that means to normalize gene length first, then sequencing depth. 
FPKM (Fragments Per Kilobase Million) takes the effect of gene lengths into consideration. These methods divide the RPM by the gene length, in kilobases. The main difference is that FPKM is made for paired-end RNA-seq, in which two reads can come from a single fragment, and FPKM can avoid counting the same fragment twice. 
```{r biomaRt, error=TRUE, tidy=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
source("../formating/formating.R")
counts2TPM <- function(count=count, efflength=efflen){   
  RPK <- count/(efflength/1000)   # Gene length normalization: reads per kilobase 
  PMSC_rpk <- sum(RPK)/1e6        # Depth normalization: “per million” scaling factor 
  RPK/PMSC_rpk                       
}
counts2FPKM <- function(count=count, efflength=efflen){    
  PMSC_counts <- sum(count)/1e6   # Depth normalization: counts “per million” scaling factor  
  FPM <- count/PMSC_counts        # Gene length normalization:  Reads/Fragments Per Million   
  FPM/(efflength/1000)                                       
} 
if("TPM" %in% methods || "FPKM" %in% methods){
  tryCatch({
    library(biomaRt)
    ensembl <- NULL
    if(params$species == "mouse") {
      ensembl = useMart("ensembl", dataset="mmusculus_gene_ensembl") # Connect to a specified BioMart database
    } else if(params$species == "human") {
      ensembl = useMart("ensembl", dataset="hsapiens_gene_ensembl") # Connect to a specified BioMart database
    } else {
      stop('You must set SPECIES to either "mouse" or "human" for TPM and FPKM normalization!')
    }
    feature_ids <- rownames(expression_matrix)
    if(params$idtype != "ENSEMBL"){
      library(AnnotationDbi)
      if(params$species == "mouse") {
        # library(EnsDb.Mmusculus.v79) 
        # ENSDB <- "EnsDb.Mmusculus.v79"
        library(org.Mm.eg.db)  
        EGDB <- "org.Mm.eg.db"
      } else if(params$species == "human") {
        # library(EnsDb.Hsapiens.v86)
        # ENSDB <- "EnsDb.Hsapiens.v86"
        library(org.Hs.eg.db)
        EGDB <- "org.Hs.eg.db"
      } else {
        stop('You must set SPECIES to either "mouse" or "human" for TPM and FPKM normalization!')
      }
      map_id <- AnnotationDbi::select(get(EGDB), keytype = params$idtype, keys =  feature_ids, columns = "ENSEMBL")
      feature_ids <- map_id$GENEID
    }
    
    attributes = c(
      "ensembl_gene_id",
      #"hgnc_symbol",
      "chromosome_name",
      "start_position",
      "end_position"
    )
    filters = "ensembl_gene_id"
    feature_info <- getBM(attributes=attributes,
                          filters=filters, 
                          values=feature_ids, mart=ensembl)
    mm <- match(feature_ids, feature_info[[filters]])
    feature_info_full <- feature_info[mm, ]
    rownames(feature_info_full) <- feature_ids
    # Caculate effective gene length
    eff_length <- abs(feature_info_full$end_position - feature_info_full$start_position)
    feature_info_full <- cbind(feature_info_full, eff_length)
    eff_length2 <- feature_info_full[,c(1,5)]
    eff_length2 <- eff_length2[order(eff_length2$ensembl_gene_id,eff_length2$eff_length,decreasing = T),] # Order by decreasing length
    g <- eff_length2[!duplicated(eff_length2$ensembl_gene_id),] # Deduplicate genes
    ng = intersect(feature_ids,g$ensembl_gene_id) 
    lengths = g[match(ng,g$ensembl_gene_id),2]
    # names(lengths) <- g[match(ng,g$ensembl_gene_id),1]
    # knitr::kable(head(lengths), align='cccc', format = "html", caption = 'Lengths')
    if("TPM" %in% methods){
      if(!"TPM" %in% names(srat@assays)){
        RedisInfo(unique_id, "Running TPM normalization.")
        tpm <- counts2TPM(expression_matrix, lengths)  
        tpm_assay <- CreateAssayObject(data = tpm)
        srat[['TPM']] <- tpm_assay
        SaveSeurat(srat, params$output)
        RedisInfo(unique_id, "Seurat object for TPM is saved successfully.")

        adata$layers["TPM"] <- t(as.data.frame(srat[['TPM']]@data))
        RedisInfo(unique_id, "TPM layer has been added to AnnData object.")
        write_h5ad(adata, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for TPM is saved successfully.")
      } else {
        RedisInfo(unique_id, "TPM assay already exists, skipped.")
        if(!"TPM" %in% names(adata$layers)){
          adata$layers["TPM"] <- t(as.data.frame(srat[['TPM']]@data))
          RedisInfo(unique_id, "TPM layer has been added to AnnData object.")
          write_h5ad(adata, adata_path, compression = "gzip")
          RedisInfo(unique_id, "AnnData object for TPM is saved successfully.")
        }
      }
      # knitr::kable(head(srat[['TPM']]@data), align='cccc', format = "html", caption = 'TPM')
      if(params$show_umap) seurat_cluster_umap(srat, assay = 'TPM', title = "TPM normalization")
    } else {
      print("Not selected by user, skipped.")
    } 
    if("FPKM" %in% methods){
      if(!"FPKM" %in% names(srat@assays)){
        RedisInfo(unique_id, "Running FPKM normalization.")
        fpkm <- counts2FPKM(expression_matrix, lengths)    
        fpkm_assay <- CreateAssayObject(data = fpkm)
        srat[['FPKM']] <- fpkm_assay
        SaveSeurat(srat, params$output)
        RedisInfo(unique_id, "Seurat object for FPKM is saved successfully.")

        adata$layers["FPKM"] <- t(as.data.frame(srat[['FPKM']]@data))
        RedisInfo(unique_id, "FPKM layer has been added to AnnData object.")
        write_h5ad(adata, adata_path, compression = "gzip")
        RedisInfo(unique_id, "AnnData object for FPKM is saved successfully.")
      } else {
        RedisInfo(unique_id, "FPKM assay already exists, skipped.")
        if(!"FPKM" %in% names(adata$layers)){
          adata$layers["FPKM"] <- t(as.data.frame(srat[['FPKM']]@data))
          RedisInfo(unique_id, "FPKM layer has been added to AnnData object.")
          write_h5ad(adata, adata_path, compression = "gzip")
          RedisInfo(unique_id, "AnnData object for FPKM is saved successfully.")
        }
      }
      # knitr::kable(head(srat[['FPKM']]@data), align='cccc', format = "html", caption = 'FPKM')
      if(params$show_umap) seurat_cluster_umap(srat, assay = 'FPKM', title = "FPKM normalization")
    } else {
      print("Not selected by user, skipped.")
    } 
  }, error = function(e) {
    RedisWarning(unique_id, paste0("An error happened during TPM/FPKM normalization normalization, skipped: ", e$message))
  })
}
```

## Gabage collection
```{r gc, error=TRUE, tidy=TRUE, include=FALSE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
srat <- NULL
adata <- NULL
expression_matrix <- NULL
```

## Session Info
```{r Session Info, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
sessionInfo()
writeLines(capture.output(sessionInfo()), gsub(".h5ad", ".txt", adata_path, ignore.case = TRUE))
```