---
title: "Quality Control using Bioconductor for `r params$dataset`"
author: "Lei Jiang"
date: "`r Sys.Date()`"
params:
  dataset: NA
  input_path: NULL # Input file path to scRNA data.
  output: NULL # Output file path of the SingleCellExperiment object.
  adata_path: NULL # Output file path of the Anndata object.
  species: "human" # c("human", "mouse") Species of the database for annotation. Allowed input is human or mouse.
  idtype: "SYMBOL" # idtype should be one of "SYMBOL", "ENSEMBL", "ENTREZID" or "REFSEQ".
  colour_by: NULL # Color by for plots
  shape_by_1: NULL # Shape by 1 for plots
  shape_by_2: NULL # Shape by 2 for plots
  output_format: SingleCellExperiment
  perplexity: 30
  n_hvg: 2000
  show_code: FALSE
  show_warning: FALSE
  show_message: FALSE
output: html_document
---
```{r setup, echo=params$show_code, include=FALSE, warning=params$show_warning, message=params$show_message}
knitr::opts_chunk$set(echo=params$show_code, include=FALSE, warning=params$show_warning, message=params$show_message)
```
## Load data
Load or create a SingleCellExperiment object from a expression matrix where each row corresponds to a gene (or transcript) and each column corresponds to a single cell.
```{r load data, include=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
library(AnnotationDbi)
library(scran)
source("../formating/formating.R")
library(RhpcBLASctl)
blas_set_num_threads(1)
omp_set_num_threads(1)
if(is.null(params$input_path)) stop("Input_path can not be NULL.")
if(is.null(params$output)) stop("Output_path can not be NULL.")
# Set altExp to contain ERCC, removing ERCC features from the main object
sce <- LoadSCE(params$input_path)
if(is.null(sce)) stop("The file format is not supported.")
altExp(sce, "ERCC") <- sce[grep("^ERCC-", rownames(sce)), ]
sce <- sce[grep("^ERCC-", rownames(sce), invert=T), ]
print("SingleCellExperiment object is loaded successfully.")
sce
```
## Annotation
Map ENSEMBL IDs to gene symbols
```{r Annotation, include=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
if(params$species == "mouse") {
  library(EnsDb.Mmusculus.v79) 
  ENSDB <- "EnsDb.Mmusculus.v79"
  library(org.Mm.eg.db)  #library(org.Hs.eg.db) if human
  EGDB <- "org.Mm.eg.db"
} else if(params$species == "human") {
  library(EnsDb.Hsapiens.v86)
  ENSDB <- "EnsDb.Hsapiens.v86"
  library(org.Hs.eg.db)
  EGDB <- "org.Hs.eg.db"
} else {
  stop('You must set SPECIES to either "mouse" or "human" at the start of this code block!')
}
if(params$idtype=="ENSEMBL"){
  gene_names <- mapIds(get(EGDB), keys=rownames(sce), keytype="ENSEMBL", columns="SYMBOL", column="SYMBOL")
  rowData(sce)$SYMBOL <- gene_names
  table(is.na(gene_names))
}
if ((params$idtype=='SYMBOL') && is.null(rowData(sce)$SYMBOL)) {
  rowData(sce)$SYMBOL <- rownames(sce)}
print("Remove all genes for which no symbols were found.")
sce <- sce[! is.na(rowData(sce)$SYMBOL),] # Remove all genes for which no symbols were found
print("Check if we can find mitochondrial proteins in the newly annotated symbols.")
grep("^MT-", rowData(sce)$SYMBOL, value=T) # Check if we can find mitochondrial proteins in the newly annotated symbols
grep("^RP[LS]", rowData(sce)$SYMBOL, value=T)
grep("ATP8", rowData(sce)$SYMBOL, value=T) # Quick search for mitochondrial protein ATP8, which is also called MT-ATP8
ensdb_genes <- genes(get(ENSDB))
MT_names <- ensdb_genes[seqnames(ensdb_genes) == "MT"]$gene_id
is_mito <- rownames(sce) %in% MT_names
table(is_mito)
```
## Basic QC
```{r Basic QC, include=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
sce_cell <- perCellQCMetrics(sce,subsets=list(Mito=is_mito))
sce_feature <- perFeatureQCMetrics(sce)
head(sce_cell)
head(sce_feature)
sce <- addPerCellQC(sce, subsets=list(Mito=is_mito))
sce <- addPerFeatureQC(sce)
hist(
    sce$total,
    breaks = 100
)
abline(v = 25000, col = "red")
hist(
  sce_cell$detected,
  breaks = 100
)
abline(v = 7000, col = "red")
```

## Identifying filtering cutoff
Adaptive threshold for filtering cutoff
```{r Identifying filtering cutoff, include=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
reasons <- quickPerCellQC(sce_cell, sub.fields=c("subsets_Mito_percent", "altexps_ERCC_percent"))
colSums(as.matrix(reasons))
sce$discard <- reasons$discard
plotColData(sce, x="sum", y="subsets_Mito_percent", colour_by="discard")
plotColData(sce, x="sum", y="detected", colour_by="discard")
plotColData(sce, x="altexps_ERCC_percent", y="subsets_Mito_percent",colour_by="discard")
library(scales)
if(!is.null(params$shape_by_1) && params$shape_by_1 %in% names(rowData(sce))){
  plotColData(sce, x="sum", y="detected", colour_by="discard", other_fields = params$shape_by_1) + 
    facet_wrap(~individual) + scale_x_continuous(labels = unit_format(unit = "k", scale = 1e-3))
}
if(!is.null(params$shape_by_2) && params$shape_by_2 %in% names(rowData(sce))){
  plotColData(sce, x="sum", y="detected", colour_by="discard", other_fields = params$shape_by_2) + 
    facet_wrap(~replicate)  + scale_x_continuous(labels = unit_format(unit = "k", scale = 1e-3))
}
```

## Highly expressed genes
The most expressed genes in the whole dataset.
```{r Highly expressed genes, include=TRUE, echo=params$show_code, warning=params$show_warning, message=params$show_message}
plotHighestExprs(sce, exprs_values = "counts", 
                 feature_names_to_plot = "SYMBOL", colour_cells_by="detected")
keep_feature <- nexprs(sce, byrow=TRUE, detection_limit = 1) >= 2
rowData(sce)$discard <- ! keep_feature
table(rowData(sce)$discard)
assay(sce, "logcounts_raw") <- log2(counts(sce) + 1) # log counts
sce.qc <- sce[! rowData(sce)$discard,! colData(sce)$discard]
# sce.qc <- sce.qc[rowSubset(sce.qc)]

# Highly variable genes
dec <- modelGeneVar(sce.qc)
chosen <- getTopHVGs(dec, n=params$n_hvg)
rowSubset(sce.qc) <- chosen # stored in the default 'subset'.

```
## Data visualization before QC
### PCA
```{r Data visualization before QC - PCA, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
if(!is.null(params$shape_by_1) && !is.null(params$colour_by) && (params$shape_by_1 %in% names(rowData(sce))) && params$colour_by %in% names(rowData(sce))){
  sce <- runPCA(sce, exprs_values = "counts")
  # dim(reducedDim(sce, "PCA"))
  plotPCA(sce, colour_by = params$colour_by, size_by = "detected", shape_by = params$shape_by_1)
  sce <- runPCA(sce, exprs_values = "logcounts_raw")
  # dim(reducedDim(sce, "PCA"))
  plotPCA(sce, colour_by = params$colour_by, size_by = "detected", shape_by = params$shape_by_1)
} else {
  sce <- runPCA(sce, exprs_values = "counts")
  # dim(reducedDim(sce, "PCA"))
  plotPCA(sce, size_by = "detected")
  sce <- runPCA(sce, exprs_values = "logcounts_raw")
  # dim(reducedDim(sce, "PCA"))
  plotPCA(sce, size_by = "detected")
}
```

### tSNE
perplexity    numeric; Perplexity parameter should not be bigger than 3 * perplexity < nrow(X) - 1
```{r Data visualization before QC - tSNE, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
set.seed(123456)
# sce <- runTSNE(sce, exprs_values = "logcounts_raw", perplexity = 130)
if(!is.null(params$perplexity)){
  perplexity <- params$perplexity
} else {
  perplexity <- floor((nrow(logcounts(sce)) - 1) / 3)
}
sce <- runTSNE(sce, exprs_values = "logcounts_raw", perplexity = perplexity)
if(!is.null(params$shape_by_1) && !is.null(params$colour_by) && (params$shape_by_1 %in% names(rowData(sce))) && params$colour_by %in% names(rowData(sce))){
  plotTSNE(sce, colour_by = params$colour_by, size_by = "detected", shape_by = params$shape_by_1)
} else{
  plotTSNE(sce, size_by = "detected")
}
```

## Data visualization after QC
### PCA
```{r Data visualization after QC - PCA, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
sce.qc <- runPCA(sce.qc, exprs_values = "logcounts_raw")
dim(reducedDim(sce.qc, "PCA"))
if(!is.null(params$shape_by_1) && !is.null(params$colour_by) && (params$shape_by_1 %in% names(rowData(sce))) && params$colour_by %in% names(rowData(sce))){
  plotPCA(sce.qc, colour_by = params$colour_by, size_by = "detected", shape_by = params$shape_by_1)
} else{
  plotPCA(sce.qc, size_by = "detected")
}
```

### tSNE
```{r Data visualization after QC - tSNE, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
set.seed(123456)
if(!is.null(params$perplexity)){
  perplexity <- params$perplexity
} else {
  perplexity <- floor((nrow(logcounts(sce.qc)) - 1) / 3)
}
sce.qc <- runTSNE(sce.qc, exprs_values = "logcounts_raw", perplexity = perplexity)
if(!is.null(params$shape_by_1) && !is.null(params$colour_by) && (params$shape_by_1 %in% names(rowData(sce))) && params$colour_by %in% names(rowData(sce))){
  plotTSNE(sce.qc, colour_by = params$colour_by, size_by = "detected", shape_by = params$shape_by_1)
} else{
  plotTSNE(sce.qc, size_by = "detected")
}
```

## Identifying Confounding Factors
### Correlations with PCs
```{r Correlations with PCs, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
sce.qc <- runPCA(sce.qc, exprs_values = "logcounts_raw")
dim(reducedDim(sce.qc, "PCA"))
if(!is.null(params$shape_by_1) && !is.null(params$colour_by) && (params$shape_by_1 %in% names(rowData(sce))) && params$colour_by %in% names(rowData(sce))){
  plotPCA(sce.qc, colour_by = params$colour_by, size_by = "sum", shape_by = params$shape_by_1)
} else{
  plotPCA(sce.qc, size_by = "sum")
}
```

### Detected genes
```{r Detected genes, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
logcounts(sce.qc) <- assay(sce.qc, "logcounts_raw")
getExplanatoryPCs(sce.qc,variables = "sum")
plotExplanatoryPCs(sce.qc,variables = "sum") 
logcounts(sce.qc) <- NULL
```

## Explanatory Variables
```{r Explanatory Variables, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
if(!is.null(params$shape_by_1) && !is.null(params$colour_by) && (params$shape_by_1 %in% names(rowData(sce))) && params$colour_by %in% names(rowData(sce))){
  plotExplanatoryVariables(sce.qc,exprs_values = "logcounts_raw",
                          variables = c("detected","sum",params$colour_by,
                                        params$shape_by_1,"altexps_ERCC_percent","subsets_Mito_percent"))
} else {
  plotExplanatoryVariables(sce.qc,exprs_values = "logcounts_raw",
                          variables = c("detected","sum","altexps_ERCC_percent","subsets_Mito_percent"))
}
```

## Save SingleCellExperiment object
```{r Save SingleCellExperiment object, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
# Drop the Alternative Experiment methods
sce.qc <- removeAltExps(sce.qc)
# Name the Main Experiment "RNA"
mainExpName(sce.qc) <- "RNA"
saveRDS(sce.qc, file = params$output)
# Convert and save SCE to AnnData object
ConvertSeuratSCEtoAnndata(params$output)
print("SingleCellExperiment object is saved successfully.")
sce <- NULL
sce.qc <- NULL
```

## Session Info
```{r Session Info, include = TRUE, echo = params$show_code, warning=params$show_warning, message=params$show_message}
sessionInfo()
writeLines(capture.output(sessionInfo()), gsub(".rds", ".txt", params$output, ignore.case = TRUE))
```